---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Gwen Rino"
date: "2/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```
***
## GITHUB LINK  

***
## THE TIDYVERSE 

```{r, echo=FALSE}
library(knitr)
table <- tibble(
  Task = c("Plotting", "Munging/wrangling", "Reshaping", "Importing/exporting"),
  Package = c("ggplot", "dplyr", "tidyr", "readr"),
  Function1 = c("geom_point", "filter", "gather", "read_delim"),
  Function2 = c("geom_smooth", "summarize", "spread", "read_csv")
)
kable(table, caption = "Name the package and two functions associated with each task")
```

***
## R BASICS  

1. Code fixed with only one change
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
```
2. Code fixed
```{r}
my_string <- c('had', 'two', 'errors', 'in', 'it')
```
3. As coded, three of the elements of `my_vector` are integers and two are characters. But vectors can only contain one type of element, so the integers are converted to characters.

***
## DATA IMPORT/EXPORT  

Demonstration of import/export:
```{r}
rail_trail <- read_delim("rail_trail.txt", delim = "|") # Load rail_trail.txt
glimpse(rail_trail)

saveRDS(rail_trail, file = "rail_trail.rds") # Export rail_trail as rds file

rail_trail_rds <- readRDS(file = "rail_trail.rds") # Reload rail_trail.rds
glimpse(rail_trail_rds)
```

***
## VISUALIZATION  

1. Three critiques of the graphic:

a. This graph requires the viewer to compare areas between the bubbles, which is difficult. This problem is somewhat mitigated by the percentages printed on the bubbles.

b. The color choices on this graph seem to be completely random. It is unclear why the age group bubbles don't have color but the gender group bubbles do.

c. If the text were bigger/bolder, it would draw attention better to the information the graph intends to convey. The names of the groups do not stand out as clearly as would be ideal.

2.
```{r}
ggplot(diamonds, aes(x=cut, y=carat)) + 
  geom_boxplot(aes(fill=color), position = "identity") + 
  xlab("CUT OF DIAMOND") + 
  ylab("CARAT OF DIAMOND") + 
  coord_flip()
```

3. It is much easier to visualize and compare the distributions of the carat variable for each combination of diamond cut and color if the graph uses the "dodge" position instead of "identity."

```{r}
ggplot(diamonds, aes(x=cut, y=carat)) + 
  geom_boxplot(aes(fill=color), position = "dodge") + 
  xlab("CUT OF DIAMOND") + 
  ylab("CARAT OF DIAMOND") + 
  coord_flip()
```

***
## DATA MUNGING AND WRANGLING  

1. The data in the column "type" needs to be spread, as shown here:
```{r}
spread(table2, key = type, value = count)
```

2. Here is the code that adds a new column to the diamonds data set called price_per_carat that shows the price of each diamond per carat.

```{r, eval=FALSE}
diamonds %>% mutate(price_per_carat = price/carat)
```

3. The output below shows how many diamonds of each cut there are total, and also how many and what proportion of these have a price > 10,000 and a carat < 1.5.

The results make sense. Only a tiny proportion of the diamonds in the dataset are both relatively small (<1.5 carats) and also relatively expensive (>10,000). However, a higher proportion of the better cut diamonds fall into this category than the poorer cut diamonds.

The number of diamonds that meet the >$10,000 & <1.5 carat requirements in each group may be misleading because the total number of diamonds in each group varies so much. But the proportion of diamonds that meet the requirements accounts for the differences in group size.

```{r}
diamonds %>% group_by(cut) %>% 
  summarize(
    total_count = n(),
    group_count = sum(price > 10000 & carat < 1.5),
    group_proportion = mean(price > 10000 & carat < 1.5)
  )
```

***
## EDA  

```{r}
txhousing
```

1. This data is from the beginning of 2000 to halfway through 2015.

```{r}
range(txhousing$date)
```

2. 46 cities are represented.

```{r}
nrow(distinct(txhousing, city))
```

3. The highest number of sales occurred in Houston in July 2015.

```{r}
txhousing %>% group_by(city, month, year) %>%
  arrange(desc(sales))
```

4. I expect listings and sales to rise and fall together.

The visualization below affirms my assumption.

```{r}
ggplot(txhousing, aes(x=listings, y=sales)) +
  geom_point(alpha=.6)
```

5. See table below for the  proportion of sales missing for each city.

```{r}
txhousing %>% group_by(city) %>% 
  summarize(proportion = mean(is.na(sales)))
```

6. Looking at only the cities and months with greater than 500 sales:

```{r}
bigsales <- txhousing %>% filter(sales > 500)
```

a. The distributions of the median sales price when grouped by city differ quite a bit. The table below shows that the city with the highest median price was Collin County ($196,000), and the city with the lowest median price was Ft. Worth ($114,600). The faceted graph below shows the distribution of median sales prices for each city. The histograms vary a lot in shape. For example, Bay Area is unimodal and nearly symmetrical, Dallas is unimodal and skews right, Austin is bimodal.

b. Any cities that stand out that you’d want to investigate further?

c. Why might we want to filter out all cities and months with sales less than 500?

```{r}
# Average median price for "big sale city-months" by city 
bigsales %>% group_by(city) %>% summarize(median = median(median)) %>% 
  arrange(desc(median)) 

# Visualize spread of median prices by city
ggplot(bigsales, aes(x = median)) + 
  geom_histogram() + 
  facet_wrap(~ city) 
```



