---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Gwen Rino"
date: "2/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

**GITHUB LINK**  

**THE TIDYVERSE**  

```{r, echo=FALSE}
library(knitr)
table <- tibble(
  Task = c("Plotting", "Munging/wrangling", "Reshaping", "Importing/exporting"),
  Package = c("ggplot", "dplyr", "tidyr", "readr"),
  Function1 = c("geom_point", "filter", "gather", "read_delim"),
  Function2 = c("geom_smooth", "summarize", "spread", "read_csv")
)
kable(table, caption = "Name the package and two functions associated with each task")
```

**R BASICS**  

1. Code fixed with only one change
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
```
2. Code fixed
```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```
3. As coded, three of the elements of `my_vector` are numbers and two are characters. But vectors can only contain one type of element, so the numbers are converted to characters.

**DATA IMPORT/EXPORT**  

Demonstration of import/export:
```{r}
rail_trail <- read_delim("rail_trail.txt", delim = "|") # Load rail_trail.txt
glimpse(rail_trail)

saveRDS(rail_trail, file = "rail_trail.rds") # Export rail_trail as rds file

rail_trail_rds <- readRDS(file = "rail_trail.rds") # Reload rail_trail.rds
glimpse(rail_trail_rds)
```

**VISUALIZATION**  

1. 

2.
```{r}
ggplot(diamonds, aes(x=cut, y=carat)) + 
  geom_boxplot(aes(fill=color), position = "identity") + 
  xlab("CUT OF DIAMOND") + 
  ylab("CARAT OF DIAMOND") + 
  coord_flip()
```

3.

**DATA MUNGING AND WRANGLING**  

1. The data in the column "type" needs to be spread, as shown here:
```{r}
spread(table2, key = type, value = count)
```

2. Here is the code that adds a new column to the diamonds data set called price_per_carat that shows the price of each diamond per carat.

```{r, eval=FALSE}
diamonds %>% mutate(price_per_carat = price/carat)
```

3. The output below shows how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5.

The results make sense. Only a tiny proportion of the diamonds in the dataset are both relatively small (<1.5 carats) and also relatively expensive (>10000). However, a higher proportion of the better cut diamonds fall into this category than the poorer cut diamonds.

Do we need to be wary of any of these numbers? Why?

```{r}
diamonds %>% group_by(cut) %>% 
  summarize(
    count = sum(price > 10000 & carat < 1.5),
    proportion = mean(price > 10000 & carat < 1.5)
  )
```

**EDA**  

```{r}
txhousing
```

1. This data is from the beginning of 2000 to halfway through 2015.

```{r}
range(txhousing$date)
```

2. 46 cities are represented.

```{r}
nrow(distinct(txhousing, city))
```

3. The highest number of sales occurred in Houston in July 2015.

```{r}
txhousing %>% group_by(city, month, year) %>%
  arrange(desc(sales))
```

4. I expect listings and sales to rise and fall together.

The visualization below affirms my assumption.

```{r}
ggplot(txhousing, aes(x=listings, y=sales)) +
  geom_point(alpha=.6)
```

5. See table below for the  proportion of sales missing for each city.

```{r}
txhousing %>% group_by(city) %>% 
  summarize(proportion = mean(is.na(sales)))
```

6. Looking at only the cities and months with greater than 500 sales:

```{r}
bigsales <- txhousing %>% filter(sales > 500)

```

Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.
Any cities that stand out that you’d want to investigate further?
Why might we want to filter out all cities and months with sales less than 500?

