---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Gwen Rino"
date: "2/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

**GITHUB LINK**

**THE TIDYVERSE**

Name the package and two functions associated with each task:
```{r, echo=FALSE}
tibble(
  Task = c("Plotting", "Munging/wrangling", "Reshaping", "Importing/exporting"),
  Package = c("ggplot", "dplyr", "tidyr", "readr"),
  Function1 = c("geom_point", "filter", "gather", "read_delim"),
  Function2 = c("geom_smooth", "summarize", "spread", "read_csv")
)
```

**R BASICS**

1. Fixed code with only one change:
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
```
2. Fixed code:
```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```
3. As entered, three of the elements of `my_vector` are numbers and two are characters. But vectors can only contain one type of element, so the numbers are converted to characters.

**DATA IMPORT/EXPORT**

Demonstration of import/export:
```{r}
rail_trail <- read_delim("rail_trail.txt", delim = "|") # Load rail_trail.txt
glimpse(rail_trail)

saveRDS(rail_trail, file = "rail_trail.rds") # Export rail_trail as rds file

rail_trail_rds <- readRDS(file = "rail_trail.rds") # Reload rail_trail.rds
glimpse(rail_trail_rds)
```

**VISUALIZATION**

1.
2.
3.

**DATA MUNGING AND WRANGLING**

1.
2.
3.

**EDA**

1.
2.
3.
4.
5.
6.

