---
title: "COMPSCIX 415.2 Homework 3"
author: "Gwen Rino"
date: "2/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

**R for Data Science Exercise 5.6.7**

2. 
```{r}
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))

# An approach that gives the same output as not_cancelled %>% count(dest)
not_cancelled %>% group_by(dest) %>% summarize(n = n()) 

# An approach that gives the same output as not_cancelled %>% count(tailnum, wt = distance)
not_cancelled %>% group_by(tailnum) %>% summarize(n = sum(distance))
```

4. It's difficult to discern a pattern for the number of cancelled flights per day. There is a spike in cancelled flights on days 39 and 40, which suggests there may have been a winter storm.

The second visualization suggests that as the average delay increases for a particular day, so does the number of cancelled flights. This makes intuitive sense!

```{r}
# Number of flights cancelled per day of year
flights %>% filter(is.na(dep_delay), is.na(arr_delay)) %>%
  group_by(month, day) %>% count(day)
# Create vector of conversions of time_hour to day of the year
doy <- as.numeric(strftime(flights$time_hour, format = "%j"))
# Mutate flights to add day_of_year variable
flights.1 <- flights %>%
  mutate(day_of_year = doy) 
# Visualize number of cancelled flights per day
cancelled <- flights.1 %>% filter(is.na(dep_delay), is.na(arr_delay)) %>%
  add_count(month, day)
ggplot(cancelled, aes(x = day_of_year, y = n)) + geom_point(alpha = .6)
# Compare proportion of cancelled flights to average delay for each day
summary <- flights.1 %>% group_by(day_of_year) %>%
  summarize(
    num_cancelled = sum(is.na(dep_delay) & is.na(arr_delay)),
    prop_cancelled = mean(is.na(dep_delay) & is.na(arr_delay)),
    ave_delay = mean(dep_delay, na.rm = TRUE)
  )
# Visualize comparison of proportion of cancelled flights to average delay
ggplot(summary, aes(x = ave_delay, y = prop_cancelled)) + geom_point(alpha = .6)
```

5. F9 is the carrier with the greatest mean dep_delay. However, F9 only flew to one destination (DEN), and only 685 flights. Other carriers different numbers of flights to different numbers of airports (for example, EV flew thousands of flights to 61 different airports). When I look at the third dataframe, sorted by mean_delay and grouped by carrier and destination, I can see that EV is the carrier that is most routinely late.

```{r}
flights %>% group_by(carrier) %>% 
  summarize(mean_delay = mean(dep_delay, na.rm = TRUE)) %>%
  arrange(mean_delay)

flights %>% group_by(carrier, dest) %>% 
  summarize(n())

flights %>% group_by(dest, carrier) %>% 
  summarize(
    mean_delay = mean(dep_delay, na.rm = TRUE),
    n = n()) %>%
  arrange(mean_delay)
```

6. The sort argument to count() sorts the output in descending order of n. It provides a quick way to see not just the max and min, but the several highest occurring counts. I can imagine many situations when this would be useful!

**R for Data Science Exercise 10.5**

Section 10.5: #1, #2, #3 and #6 only

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

6. What option controls how many additional column names are printed at the footer of a tibble?

**R for Data Science Exercise 12.3.3**

2. Why does this code fail?

table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in combine_vars(vars, ind_list): Position must be between 0 and n

3. Why does spreading this tibble fail? How could you add a new column to fix the problem?

people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?

preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

**R for Data Science Exercise 12.4.3**

1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))

2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

**Final Question**

```{r}
baby_names <- read_delim("baby_names.txt", delim = "|") # Load baby_names.txt
glimpse(baby_names)

saveRDS(baby_names, file = "baby_names.rds") # Export baby_names as rds file

baby_names_rds <- readRDS(file = "baby_names.rds") # Reload baby_names.rds
glimpse(baby_names_rds)
```

