---
title: "Homework_2_Rino_Gwen"
author: "Gwen Rino"
date: "2/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(gridExtra)
```

**R for Data Science Exercise 3.2.4**

1. The code `ggplot(data = mpg)` creates just an empty gray box.

2. Using `glimpse(mpg)`, I can see that there are 234 rows and 11 columns in the mpg dataset.

3. `drv` is a categorical variable with three levels: f = front-wheel drive, r = rear wheel drive, 4 = 4wd

4. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))
```

5. The plot below is not useful because both variables are categorical.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = class))
```

**R for Data Science Exercise 3.3.1**

1. The code below is wrong because "color" is not mapped to a variable and so should not be inside the `aes()`.

```{r, eval = FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

2. You can see the variable type at the top of each column when you run `mpg`.

Categorical variables: manufacturer, model, year, cyl, trans, drv, fl, class  
Continuous variables: displ, cty, hwy

3. "Color" uses different gradients of the same color for continuous variables, and completely different colors for each level of a categorical variable. "Size" uses size gradients for continuous variables, while it assigns a different size point for each level of a categorical variable. "Shape" doesn't work at all with continuous variables (which makes sense!).

4. When you map the same variable to multiple aesthetics, the variable is identified by more than one quality. In the plot below, each point's fuel type is shown by both shape and color.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = fl, color = fl))
```

5. "Stroke" modifies the width of the border of shapes that have borders, as shown below.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = cyl), shape = 21, color = "black", fill = "white", size = 5, stroke = 1)
```

6. This aesthetic `aes(color = displ < 5)` creates a plot in which the value of the displ variable (greater or less than 5) is distinguished by color, as shown below.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, color = displ < 5))
```

**R for Data Science Exercise 3.5.1**

4. Advantages of using faceting: The faceted variable really stands out as the strongest element of the visualization. For example, this plot invites the viewer to compare the engine size and highway mileage for each class of car. The class of car is the primary visual component. The plot answers the question, how do different classes of car compare on engine size and highway mileage?
```{r, echo = FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Advantages of using color: The x and y variables are the strongest element of a visualization that adds a third variable with color. For example, this plot invites the viewer to see the negative correlation between engine size and highway mileage for cars overall. The class of car is a secondary point of information, more of an impression generated by the color. The plot answers the question, how are car engine size and highway mileage correlated? It also invites a second question, is class of car relevant to the engine size/highway mileage correlation?
```{r, echo = FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Considerations for a larger dataset: If there are many levels of the third variable, then color is probably a better choice than faceting. If there are many observations in the dataset but not too many levels of the third variable, then faceting would be the better choice. Most of all, though, I think the decision depends on how much you want to emphasize the third variable vs. the relationship between x and y.

5. `nrow` determines the number of rows and `ncol` determines the number of columns the the layout of the panels of a faceted plot. You can also change how the labels are printed, the order in which the panels appear, and the scale of the axes of the panels.

`facet_grid()` doesn't have `nrow` and `ncol` arguments because the arrangement of the panels is determined by the combinations of levels of the two variables on which the plot is faceted.

**R for Data Science Exercise 3.6.1**

1. Line chart: `geom_line()`  
Boxplot: `geom_boxplot()`  
Histogram: `geom_histogram()`  
Area chart: `geom_area()`

2. I imagine this visualization will plot `displ` against `hwy` on a scatter plot with `drv` levels distinguished by color. There will also be a smooth curve best fit line. (I was right about everything except that there is a separate smooth curve for each level of `drv`. Of course this makes sense because `drv` is mapped as a global aesthetic in the `ggplot()` line of code, so every geom will include this aesthetic.)
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3. `show.legend = FALSE` means that the legend for a particular geom will not be displayed. If you remove this code, the default is that the legend will be displayed. I think the author used `show.legend = FALSE` earlier in the chapter so that the plot with color mapped to `drv` would be parallel to the other two plots in the example, neither of which has an automatically-generated legend.

4. The `se` argument shows the confidence interval around the curve as a gray shadow. The default is `se = TRUE`.

5. These two plots are identical. In the first version the data and aesthetics are included in the `ggplot()` line, so they are global mappings that are applied to every geom layer. In the second version the mappings are applied at each geom layer. Because the mappings are identical for each geom layer, this version is less efficient than the first version.
```{r}
p1 <- ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

p2 <- ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

grid.arrange(p1, p2, ncol = 2)
```

**R for Data Science Exercise 3.7.1**

2. `geom_bar()` plots just one variable, showing the counts for each level as a bar graph. The height of each bar is proportional to the number of cases in each group. `geom_col()` plots a categorical variable x against a continuous variable y and shows the results as a bar graph. The height of the bars represents values in the data.

```{r}
p3 <- ggplot(mpg, aes(fl)) + geom_bar()

p4 <- ggplot(mpg, aes(class, hwy)) + geom_col()

grid.arrange(p3, p4, ncol = 2)
```

**Final Question**

Evaluation of data graphics at [What is a Data Scientist?](https://onlinelearning.berkeley.edu/courses/882865/files/folder/Session%202%20-%20Feb%206%202018?preview=60649489)

The first graphic, a pie chart showing somebody's beliefs (we don't know whose) about the demand for data scientists, is not a very effective visualization. The point, that there is high demand for data scientists, is not easy to discern. I would suggest a clearer title and a different kind of plot that makes the high demand for data scientists more apparent.

The second graphic, a bar graph showing the sources of new data talent, is much more effective. It is easy to read and interpret. I like that the bars are proportionally sized in width as well as height.

The third graphic, a chart showing the percentage of organizations that identify various obstacles to adopting data science, is reasonably effective. I like that the visualization includes progressive transparency for smaller and smaller percentages.

The fourth graphic is a faceted bar graph showing the level of education of data scientists compared to business intelligence professionals. I would suggest a much clearer title in order to increase the effectiveness of this visualization.

The fifth graphic uses proportionally-sized bubbles to show the areas of study of data scientists compared with business intelligence professionals. I think that this graphic is meant to emphasize in a positive way that data scientists have a wider range of academic backgrounds than BI professionals do, but when I look at the graphic my eye is drawn to the largest bubble, which denotes that 37% of BI professionals have a degree in business. Because my eye is drawn there, I am inclined to think that this means something positive. I don't think the bubbles are the best means of visualizing the (desirable) diversity of backgrounds of data scientists compared with BI professionals.

The sixth graphic is a pie chart showing the proportion of data scientists who believe that their work makes more work for data scientists vs. those who do not believe this. I am interested to notice that a pie chart with only two levels is much easier to interpret than a pie chart with many levels. That said, I think that the title and labels of this visualization are a little confusing and could be improved.

The seventh graphic is a T-chart comparing the qualities of a "big data scientist" and a "normal data scientist". In my opinion, "normal data scientist" is an unfortunate choice of title (since the opposite of "normal" is "abnormal", and who wants to be associated with abnormality?). There is also a proportional bar below the T-chart showing (I think!) the percentage of data scientists who work in big data, "normal" data, or somewhere in the middle. Another possible meaning of this proportional bar is that a typical data scientist works 10% of the time in big data, 25% of the time in "normal" data science, and 65% of the time in the middle. I have to guess because there is no title or legend explaining this part of the visualization. I do like it that the proportional bar is organized in space in a meaningful way (with big data and "normal" data to the same sides as the T-chart, and "middle" in between).

The eighth graphic is a faceted bar graph, like the fourth graphic, showing the percentage of data scientists who work in various aspects of the data cycle compared with the percentage of BI professionals who do. It takes very careful inspection to understand that the blue bars represent BI professionals, and this is especially confusing because in the previous visualization blue was associated with "normal" data scientists. I would recommend that the colors be used consistently to represent the same group throughout. I would also recommend a clearer legend for this graphic.

The ninth and final graphic is extremely confusing. There are two titles, "Who does a data scientist work with?" and "Data science team." There are a number of job titles and associated percentages. But the meaning of the numbers is impossible to discern. Do 36% of data science teams include a programmer? Do 36% of data scientists work with a programmer? Are 36% of the people on data science teams programmers? It is impossible to tell. This visualization needs to be completely overhauled in order to tell a clear story.