---
title: "COMPSCIX 415.2 Homework 3"
author: "Gwen Rino"
date: "2/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

**R for Data Science Exercise 3.8.1**

1. The problem with the graph is that it is overplotted (points overlap each other). It could be improved by using `position = "jitter"` to add random noise that will spread the points out a bit and make them easier to see and interpret.

2. The parameters `width` and `height` control the amount of jitter.

3. `geom_jitter()` and `geom_count()` are two different ways of dealing with overplotting. `geom_jitter()` adds random noise to spread out the points, while `geom_count()` counts the number of observations at each location, then maps the count to the size of the point.

4. The default position adjustment for `geom_boxplot()` is dodged. For example:

```{r}
ggplot(mpg, aes(class, hwy)) + 
  geom_boxplot(aes(color = drv))
```

**R for Data Science Exercise 3.9.1**

2. `labs()` is used to modify labels of plot axes, legends, titles, subtitles, and captions.

4. The plot shows several things: there is a positive linear relationship between city mpg and highway mpg, all cars are more fuel efficient on the highway than in the city, and there is (generally) a greater difference between city and highway fuel efficiency for cars that get higher mpg in the city (that is, the slope of the best fit line is greater than 1). `coord_fixed()` makes the scale for each axis the same, which is important for visualizing the increasing difference between city and highway fuel efficiency as city fuel efficiency increases. `geom_abline` adds the reference diagonal with a slope of 1, which further emphasizes this same relationship.

**R for Data Science Exercise 4.4**

1. The code doesn't work because the character "i" has been replaced with a "1".

2. Corrected code:

```{r, eval = FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)

filter(diamonds, carat > 3)
```

**R for Data Science Exercise 5.2.4**

1. Find all flights that:

a. Had an arrival delay of two or more hours

```{r, eval = FALSE}
filter(flights, arr_delay >= 120)
```

b. Flew to Houston (IAH or HOU)

```{r, eval = FALSE}
filter(flights, dest == "IAH" | dest == "HOU")
# OR
filter(flights, dest %in% c("IAH", "HOU"))
```

c. Were operated by United, American, or Delta

```{r, eval = FALSE}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

d. Departed in summer (July, August, and September)

```{r, eval = FALSE}
filter(flights, month %in% c(7,8,9))
```

e. Arrived more than two hours late, but didn’t leave late

```{r, eval = FALSE}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

f. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r, eval = FALSE}
filter(flights, dep_delay >= 60 & arr_delay < (dep_delay - 30))
```

g. Departed between midnight and 6am (inclusive)

```{r, eval = FALSE}
filter(flights, dep_time <= 600)
```

2. `between()` tests whether values in a numeric vector fall within a specified range. It could be used as alternatives as noted here:

```{r, eval = FALSE}
flights[between(flights$month, 7, 9), ] # 1.d.
flights[between(flights$dep_time, 0, 600), ] # 1.g.
```

3. 8,255 rows have a missing `dep_time`, and these rows are also missing `dep_delay`, `arr_time`, `arr_delay`, and `air_time`. It seems likely that these flights were canceled.

```{r}
filter(flights, is.na(dep_time))
```

4. 
```{r}
NA ^ 0 # Any value, even an unknown, raised to the 0 power = 1.

NA | TRUE # If either part of a Boolean OR is TRUE, the whole statement is TRUE. Doesn't matter if the other part is unknown.

FALSE & NA # If either part of a Boolean AND is FALSE, the whole statement is FALSE. Doesn't matter if the other part is unknown.

NA * 0 # Well, this IS a tricky counterexample! Seems like the answer should be 0, because any value, even an unknown, multiplied by 0 = 0.
```

**R for Data Science Exercise 5.4.1**

1. Here are a number of ways to select dep_time, dep_delay, arr_time, and arr_delay from flights:

```{r, eval = FALSE}
select(flights, dep_time, dep_delay, arr_time, arr_delay)

select(flights, -year, -month, -day, -sched_dep_time, -sched_arr_time,
       -carrier, -flight, -tailnum, -origin, -dest, -air_time, -distance,
       -hour, -minute, -time_hour)

select(flights, starts_with("dep"), starts_with("arr"))

select(flights, 4, 6, 7, 9)

select(flights, 4:9, -5, -8)

select(flights, dep_time:arr_delay, -sched_dep_time, -sched_arr_time)

select(flights, dep_time:arr_delay, -(starts_with("sched")))

select(flights, starts_with("dep"), 7, 9)

select(flights, starts_with("arr"), 4, 6)

select(flights, ends_with("time"), ends_with("delay"), -sched_dep_time,
       -sched_arr_time, -air_time)

select(flights, ends_with("delay"), dep_time, arr_time)

```

2. Even if you include the name of a variable multiple times in a `select()` call, the variable is only selected once.

3. If you create a vector of the names of the variables of interest, then you can use `one_of()` with `select()` to select the variables in the vector. See the example below.

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

4. The select helpers ignore case by default. If you want the helper to be case-sensitive, add the argument `ignore.case = FALSE`