---
title: "COMPSCIX 415.2 Homework 8"
author: "Gwen Rino"
date: "3/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(rpart)
library(partykit)
library(ROCR)
```

***
## EXERCISE ONE  

This data set has 891 observations and 12 columns.  

```{r}
titanic <- read_csv("Titanic.csv")
nrow(titanic)
ncol(titanic)
```

***
## EXERCISE TWO  

```{r}
# Set seed
set.seed(29283)

# Convert Survived and Pclass to factor
titanic$Pclass <- factor(titanic$Pclass)
titanic$Survived <- factor(titanic$Survived)

# Create training set
train_set <- titanic %>% sample_frac(0.7)

# Create test set (observations in titanic$PassengerId that are not in train_set$PassengerId)
test_set <- titanic %>% filter(!(titanic$PassengerId %in% train_set$PassengerId))
```

***
## EXERCISE THREE  

Preliminary to fitting the model...  
* My guess is that Pclass level 2 and 3 are associated with a lower probability of survival than Pclass level 1, Sex (female) is associated with a higher probability of survival ("Women and children first!"), and higher Fares are associated with higher probability of survival.  
* My concern about this choice of features is that Fare and Pclass are probably correlated.  

```{r}
# Fit the logistic regression model
log_mod.1 <- glm(Survived ~ Pclass + Sex + Fare, data = train_set, family = "binomial")
# Output coefficients
tidy(log_mod.1)
```

The coefficients in the model show that:  
* The lower class the passenger is, the lower the probability of survival.  
* Being male is associated with a lower probability of survival.  
* Paying a higher fare is associated with a higher probability of survival.  

The p-values of the coefficients show that Pclass and Sex are statistically significant predictors of survival, but Fare is not.

***
## EXERCISE FOUR

a. A third class female passenger who paid a fare >= 23.7 has a low probability of survival.  
b. I'm surprised that the tree shows such different probabilities for survival among third class female passengers who paid different fares. Higher fares do not seem to be associated with higher probability of survival among third class female passengers.  
```{r}
tree_mod.1 <- rpart(Survived ~ Pclass + Sex + Fare, data = train_set)
plot(as.party(tree_mod.1))
```

***
## EXERCISE FIVE  


