---
title: "Homework_3_Rino_Gwen"
author: "Gwen Rino"
date: "2/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

**R for Data Science Exercise 3.8.1**

1. The problem with this graph is that it is overplotted (points overlap each other). It could be improved by inserting `position = "jitter"` to add random noise that will spread the points out a bit and make them easier to see and interpret.

2. The parameters `width` and `height` control the amount of jitter.

3. `geom_jitter()` and `geom_count()` are two different ways of dealing with overplotting. `geom_jitter()` adds random noise to spread out the points, while `geom_count()` counts the number of observations at each location, then maps the count to the size of the point.

4. The default position adjustment for `geom_boxplot()` is dodged. For example:

```{r}
ggplot(mpg, aes(class, hwy)) + 
  geom_boxplot(aes(color = drv))
```

**R for Data Science Exercise 3.9.1**

2. `labs()` is used to modify labels of plot axes, legends, titles, subtitles, and captions.

4. This plot shows several things: there is a positive linear relationship between city mpg and highway mpg, all cars are more fuel efficient on the highway than in the city, and there is (generally) a greater difference between city and highway fuel efficiency for cars that get higher mpg in the city.

`coord_fixed()` makes the scale for each axis the same, which is important for visualizing the increasing difference between city and highway fuel efficiency as city fuel efficiency increases. `geom_abline` adds the reference diagonal with a slope of 1, which further emphasizes this same relationship.

**R for Data Science Exercise 4.4**

1.

2.

**R for Data Science Exercise 5.2.4**

1.

2.

3.

4.

**R for Data Science Exercise 5.4.1**

1.

2.

3.

4.